Discovering the Rules of language using Bayesian Program Learning

### 2018-04-29 outline.tex
heuristic description

Combining \texttt{solveIncrementalChange} with CEGIS gives the
skeleton of the inference algorithm that I have been experimenting
with, called \texttt{incrementalCEGIS}. See the pseudocode at the top of the next page:
\pagebreak
\begin{algorithmic}
  \STATE \textbf{\texttt{incrementalCEGIS}}:
  \STATE {\bfseries Input:} Matrix of inflected forms for $N$ lexemes, $X$
  \STATE {\bfseries Output:} underlying forms, $\{u_n\}$; prefixes $\{p_i\}$; suffixes $\{s_i\}$; rules $\{r_j\}$
  \STATE // Initialize training set $T$ with a single example
  \STATE Set $T = $ inflections for any lexeme in $X$
  \STATE // Find a solution for this single example, using as many rules as you need ($\infty$)
  \STATE Set $\{r_j\}_{j = 1}^{B_R}, \{p_i\}, \{s_i\} = \text{solver}(T,\infty)$
  \REPEAT
  \STATE // Find the set of counterexamples $C$ inconsistent with our current guess as to what the solution is
  \STATE Set $C = \{x | x\in X \text{ and } x \text{ inconsistent with } \{r_j\}_{j = 1}^{B_R},\{p_i\}, \{s_i\}\}$
  \STATE // If there are no counterexamples as we are done
  \STATE \textbf{if} $C = \varnothing$ \textbf{ then return }$\{r_j\}_{j = 1}^{B_R},\{p_i\}, \{s_i\}$

  \STATE // Start out with a search radius of $1$ and increase it until we find a way of accommodating another example
  \FOR{$R = 1,2,3,4,...$}
  \FOR {$c\in C$}
  \STATE Set synthesisResult$ = \text{solveIncrementalChange}(T\cup\{c\},\{r_j\},R)$
  \IF{synthesisResult$\not= \bot$}
  \STATE Set $\{r_j\}_{j = 1}^{B_R}, \{p_i\}, \{s_i\} = $synthesisResult
  \STATE Add $c$ to the set of training examples, $T$
  \STATE \textbf{break out of loop over }$R$ // Successfully updated both our solution and added another example to the training set
\end{algorithmic}
\vspace{1cm}

@@

>>>>

Combining \texttt{solveIncrementalChange} with CEGIS gives the
skeleton of the inference algorithm that I have been experimenting
with, called \texttt{incrementalCEGIS}. See the pseudocode at the top of the next page:
\pagebreak
\begin{algorithmic}
  \STATE \textbf{\texttt{incrementalCEGIS}}:
  \STATE {\bfseries Input:} Matrix of inflected forms for $N$ lexemes, $X$
  \STATE {\bfseries Hyperparameter:} Batch size $B$
  \STATE {\bfseries Output:} underlying forms, $\{u_n\}$; prefixes $\{p_i\}$; suffixes $\{s_i\}$; rules $\{r_j\}$
  \STATE // Initialize training set $T$ with $B$ examples
  \STATE Set $T = $ inflections for $B$ lexemes in $X$
  \STATE // Find a solution for examples, using as many rules as you need ($\infty$)
  \STATE Set $\{r_j\}_{j = 1}^{B_R}, \{p_i\}, \{s_i\} = \text{solver}(T,\infty)$
  \REPEAT
  \STATE // Find the set of counterexamples $C$ inconsistent with our current guess as to what the solution is
  \STATE Set $C = \{x | x\in X \text{ and } x \text{ inconsistent with } \{r_j\}_{j = 1}^{B_R},\{p_i\}, \{s_i\}\}$
  \STATE // If there are no counterexamples as we are done
  \STATE \textbf{if} $C = \varnothing$ \textbf{ then return }$\{r_j\}_{j = 1}^{B_R},\{p_i\}, \{s_i\}$
  \STATE // Add a batch of counterexamples to the training set
  \STATE $T\gets T\cup \{C_j\}_{j = 1}^B$
  \STATE // Start out with a search radius of $1$ and increase it until we find a way of accommodating another example
  \FOR{$R = 1,2,3,4,...$}
%  \FOR {$c\in C$}
  \STATE Set synthesisResult$ = \text{solveIncrementalChange}(T,\{r_j\},R)$
  \IF{synthesisResult$\not= \bot$}
  \STATE Set $\{r_j\}_{j = 1}^{B_R}, \{p_i\}, \{s_i\} = $synthesisResult
  \STATE \textbf{break out of loop over }$R$ // Successfully updated both our solution and the training set
\end{algorithmic}
\vspace{1cm}

@@

\subsection{Search heuristics}

\subsubsection{Guessing underlying forms}

\textbf{incrementalCEGIS} is a generic way of learning large programs that explain large data sets,
and does not encode any domain specific knowledge about phonology.
Empirically, the search problem is \emph{much} easier if you
either know what the underlying forms are or have some constraints on what the underlying forms are.
If I introspect on my own reasoning process while solving these problems,
I usually guess at most a few plausible underlying forms for each lexeme.
I encoded a simple heuristic for guessing parts of underlying representations:
given prefixes/suffixes $\left\{p_i \right\}$ \& $\left\{s_i \right\}$
and observed inflections for a single lexeme $\left\{x_i \right\}$,
this heuristic guesses that the underlying form begins with the longest shared prefix
of the surface forms with the prefixes chopped off,
and that it ends with the longest shared suffix of the surface forms with the suffixes chopped off.
For example, if we have surface forms $(\text{test},\text{testing},\text{tests})$,
and have prefixes $(\text{t},\text{t},\text{t})$ and suffixes $(\text{t},\text{ting},\text{ts})$,
this heuristic constrains the underlying form to start with $\text{es}$ and end with $\text{es}$.
